# Lasso
Implementing GLASSO in a Gaussian Tree Model

---
title: "Glasso on a Gaussian Tree"
---

1. Generate A random Sample for a multivariate Gaussian density with p=11 and mu=0. The covariance matrix has normalized diagonal entries. Also, the covariance matrix Sigma should be related to a Tree with 11 nodes, that is drawn in the report: 

```{r}
library('MASS')
#Sigma <- matrix(c(10,2,1,1,4,4,2,10,0.2,0.2,0.8,0.8,1,0.2,10,0.1,0.4,0.4,1,0.2,0.1,10,0.4,0.4,4,0.8,0.4,0.4,10,1.6,4,0.8,0.4,0.4,1.6,10),nrow=6)
Sigma<-rbind(c(1,0.6,0.42,0.18,0.027,0.144,-0.0288,-0.1008,0.036,-0.084,0.168),
             c(0.6,1,0.7,0.3,0.045,0.24,-0.048,-0.168,0.06,-0.14,0.28),
             c(0.42,0.7,1,0.21,0.0315,0.168,-0.0336,-0.1176,0.042,-0.2,0.4),
             c(0.18,0.3,0.21,1,0.15,0.8,-0.16,-0.56,0.2,-0.042,0.084),
             c(.027,.045,.0315,.15,1,.12,-.024,-.084,.03,-.0063,.0126),
             c(0.144,0.24,0.168,0.8,0.12,1,-0.2,-0.7,0.25,-0.0336,0.0672),
             c(-.0288,-.048,-.0336,-.16,-.024,-.2,1,.14,-.05,.00672,-0.1344),
             c(-.1008,-.168,-.1176,-.56,-.084,-.7,.14,1,-.175,.02352,-.04704),
             c(0.036,0.06,0.042,0.2,0.03,0.25,-0.05,-0.175,1,-0.0084,0.0168),
             c(-.084,-.14,-.2,-.042,-.0063,-.0336,.00672,.02352,-.0084,1,-.08),
             c(.168,.28,.4,.084,.0126,.0672,-.1344,-.04704,.0168,-.08,1)       
             )
print(Sigma)
mu <- rep(0, times=11)
samples <- mvrnorm(n = 100, mu, Sigma, tol = 1e-6, empirical = FALSE, EISPACK = FALSE)
#print(samples)
```

2. It is recommended to standardize our sample data (each column) (but for a moment we ignore this constraint), and compute the empirical covariance matrix S. We also plot the ML precision matrix (note that all the paths regarding to plot functions should be changed appropriatley and set to the local path on the corresponding machine.)

```{r}
#scaledsamples<-scale(samples)
#S<-cov(scaledsamples)
S <- cov(samples)
S2<-S
colnames(S2) <- c("X1","X2","X3","X4","X5","X6","X7","X8","X9","X10","X11")
rownames(S2) <- c("X1","X2","X3","X4","X5","X6","X7","X8","X9","X10","X11")
ML.precision<-solve(S2) # gives the ML precision matrix
#######################################################################
library('gplots')
png("D:/LSU/Spring 2015/Detection and Estimation/Project/R codes/ML.precision.png",    # create PNG for the heat map        
width = 5*300,        # 5 x 300 pixels
height = 5*300,
res = 300,            # 300 pixels per inch
pointsize = 8)        # smaller font size
######################################################################
my_palette <- colorRampPalette(c("black", "white", "black"))(n = 200)
par(mfrow=c(2,1))
heatmap.2(ML.precision,
  main = "ML Precision", # heat map title
  notecol="black",      # change font color of cell labels to black
  density.info="none",  # turns off density plot inside color legend
  trace="none",         # turns off trace lines inside the heat map
  Colv="NA",            # turns off column clustering
  Rowv="NA",
  col=my_palette,
  dendrogram = "none"
  )
dev.off()
```

3. Now, we find the precision matrix using graphical lasso:

```{r}
library('glasso')
estimated<-glasso(S,rho=0.1) # rho is the normalized version of lambda
print(estimated$wi) # precision using Glasso
print(solve(S2)) # Finds the inverse of S2
print(solve(Sigma))
#print(estimated$w)  # covariance using Glasso
```

4. Let's write the precision matrix into Data

```{r}
write.csv(
estimated$wi,"D:/LSU/Spring 2015/Detection and Estimation/Project/R codes/Lasso.precision.csv",row.names=FALSE)

```

5. Let's plot the heatmap for actual precision matrix

```{r}

colnames(Sigma) <- c("X1","X2","X3","X4","X5","X6","X7","X8","X9","X10","X11")
rownames(Sigma) <- c("X1","X2","X3","X4","X5","X6","X7","X8","X9","X10","X11")
REAL.precision<-solve(Sigma) # gives the REAL precision matrix
#######################################################################
#library('gplots')
png("D:/LSU/Spring 2015/Detection and Estimation/Project/R codes/REAL.precision.png",    # create PNG for the heat map        
width = 5*300,        # 5 x 300 pixels
height = 5*300,
res = 300,            # 300 pixels per inch
pointsize = 8)        # smaller font size
######################################################################
my_palette <- colorRampPalette(c("black", "white", "black"))(n = 200)
#par(mfrow=c(2,1))
heatmap.2(REAL.precision,
  main = "REAL Precision", # heat map title
  notecol="black",      # change font color of cell labels to black
  density.info="none",  # turns off density plot inside color legend
  trace="none",         # turns off trace lines inside the heat map
  Colv="NA",            # turns off column clustering
  Rowv="NA",
  col=my_palette,
  dendrogram = "none"
  )
dev.off()
```


6. Let's plot the LASSO precision matrix

```{r}

LASSO.precision<-estimated$wi
LASSO.cov<-estimated$w
colnames(LASSO.precision) <- c("X1","X2","X3","X4","X5","X6","X7","X8","X9","X10","X11")
rownames(LASSO.precision) <- c("X1","X2","X3","X4","X5","X6","X7","X8","X9","X10","X11")
#REAL.precision<-solve(Sigma) # gives the REAL precision matrix
#######################################################################
#library('gplots')
png("D:/LSU/Spring 2015/Detection and Estimation/Project/R codes/LASSO.precision.png",    # create PNG for the heat map        
width = 5*300,        # 5 x 300 pixels
height = 5*300,
res = 300,            # 300 pixels per inch
pointsize = 8)        # smaller font size
######################################################################
my_palette <- colorRampPalette(c("black", "white", "black"))(n = 200)
#par(mfrow=c(2,1))
heatmap.2(LASSO.precision,
  main = "LASSO Precision", # heat map title
  notecol="black",      # change font color of cell labels to black
  density.info="none",  # turns off density plot inside color legend
  trace="none",         # turns off trace lines inside the heat map
  Colv="NA",            # turns off column clustering
  Rowv="NA",
  col=my_palette,
  dendrogram = "none"
  )
dev.off()

```


Let's Find the KL-distance between KL[Lasso||real] and KL[ML||real]

```{r}
library('matrixcalc')
KL.lasso.real<-0.5*(matrix.trace((solve(Sigma))%*%(LASSO.cov))+log(det(Sigma)/det(LASSO.cov)))
##############################################################################
KL.ML.real<-0.5*(matrix.trace((solve(Sigma))%*%(S))+log(det(Sigma)/det(S)))
##############################################################
print(KL.lasso.real)
print(KL.ML.real)
#print(LASSO.cov)
#print(S)
#print(LASSO.cov-S)
```


Let's plot the KL distances (the vectors kl.lasso.real and kl.ml.real are found by several rounds of simulations and taking the expected values. The more appropriate way should be a more exhaustive rounds of simulations (more than hundreds), since more exhaustive simulations takes more time, here I used only several rounds of simulations):

```{r}
png("D:/LSU/Spring 2015/Detection and Estimation/Project/R codes/KL.png",    # create PNG for the heat map        
width = 5*300,        # 5 x 300 pixels
height = 5*300,
res = 300,            # 300 pixels per inch
pointsize = 8) 
#########################################################################
########################################################################
kl.lasso.real <- c(5.958,5.945,5.954,5.98,5.99,6.045,6.09,6.07,6.1,6.08,6.1,6.06,6.23,6.23,6.33)
kl.ml.real <- c(5.5165,5.525,5.54,5.57,5.61,5.82,5.89,6.08,6.12,6.06,6.22,6.42,6.96,7.43,8.05)
N.real <- c(2000,1500,1000,500,300,100,90,80,70,60,50,40,30,20,15)
N<-c(1:15)
plot(N,kl.lasso.real,type="b",col="red",ylim=range(c(kl.lasso.real,kl.ml.real)),ann=FALSE,xaxt='n')
par(new=TRUE)
plot(N,kl.ml.real,type="b",col="blue",ylim=range(c(kl.lasso.real,kl.ml.real)),axes=FALSE,main="Performance Analysis of LASSO vs. ML",xlab="Number of Samples(N)",ylab="KL Distance: KL(LASSO/ML||True Covariance)")
box() # includes box around the plot
axis(1, at=1:15, lab=N.real)
legend("toplef",legend=c("KL[LASSO||Real]","KL[ML||REAL]"),col=c("red","blue"),lty=c(1,1))
dev.off()
```
